### 2011-06-08 paper/end.tex
Results table now in implementation tex, but correctly shows up in results section.

\def\topfraction{.9}
\def\floatpagefraction{.8}

\section{Results}

\section{Conclusions}
ending stuff